---
title: "p8105_HW3"
output: github_document
date: "2023-10-13"
---

```{r setuo, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

## Problem 1

### load the dataset

```{r}
library(p8105.datasets)
data("instacart")
instacart
```

### Numbr of aisles and most items ordered from
```{r}
instacart |>
  count(aisle) |>
  arrange(desc(n))
```

#### There are 134 aisles and fresh vegetables is the aisle that the most items ordered from 

### Plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered
```{r}
instacart |>
  count(aisle) |>
  filter(n > 10000) |> 
  ggplot(aes(x = reorder(aisle, -n), y = n)) + geom_point() + labs(title = "Number of items ordered each aisle", x = "Aisle Name", y = "Number of items ordered") +       theme(axis.text.x = element_text(angle = 50, hjust = 1))

```
### Making a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care",  "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |>
  mutate(rank = min_rank(desc(n))) |>
  filter(rank < 4) |>
  arrange(desc(n)) |>
  knitr::kable()
```

### Making a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```





